---
title: "dplyr"
format: html
editor: visual
---

```{r, message=FALSE}
library(tidyverse)
library(DBI)
library(duckdb)
library(dbplyr)
```
Question: 
- Is collect(), us fetching for the database?

## Connecting to a database
```{r}
conn <- DBI::dbConnect(duckdb::duckdb(),
                       # Connect to database on drive
                       dbdir = "database/database.duckdb")

# Check if connection worked
conn
```
# Check if connected to database - Check for tables in the database
### Same thing as `.table` in python
```{r}
DBI::dbListTables(conn)
```

# Look at one table
It returns another object- It doesn't know how many rows it has
```{r}
species <- tbl(conn, 
               # Select table
               "Species")
# Look at table
species

# Simliar to this
# species <- read_csv()
```

# Let's do some queries

Question: Is this selecting from the whole database or only from when we looked at it from the code chunk above?

```{r}
species_a <- dbGetQuery(conn, 
                        # Pass Query
                        "SELECT * FROM Species WHERE Code LIKE 'a%'") # No semi colon

species_a


```

## Dbplyer
```{r}
species %>% 
  filter(Relevance == "Study species") %>% 
  select(Code, Scientific_name) %>% 
  arrange(Scientific_name) %>% 
  head(3) %>% 
  # Could get SQL Code for this 
  show_query()
  
```

```{r}
species %>% 
  filter(Relevance == "Study species") %>% 
  select(Code, Scientific_name) %>% 
  arrange(Scientific_name) %>% 
  head(3) %>% 
  # get the entire database
  collect()
  
```


# Lets do a group_by, summarize
```{r}
species %>% 
  group_by(Relevance) %>% 
  summarize(num_species = n()) %>% 
  #show_query()
  collect()
```

# Mutate
```{r}
species_x <- species %>% 
  mutate(Xcode = paste("X_", Code)) %>% 
  #show_query()
  collect()
```
# Let's do a Join
To start off we will load in the tables
```{r}
eggs_db <- tbl(conn, "Bird_eggs")

nest_db <- tbl(conn, "Bird_nests")
```

# 
```{r}
nest_eggs_df <- nest_db %>% 
  left_join(eggs_db, by = "Nest_ID", keep=FALSE) %>% 
  head(5) %>% 
  #show_query() # Not the best way to do an SQL for joining
  collect()
```

# Closing Connnection
Important to do when finishing
```{r}
DBI::dbDisconnect(conn, shutdown = TRUE)
```




