---
title: "hw6_1"
format: html
---

```{r, message=FALSE}
library(tidyverse)
library(DBI)
library(duckdb)
library(dbplyr)
```


## Connecting to a database
```{r}
conn <- DBI::dbConnect(duckdb::duckdb(),
                       # Connect to database on drive
                       dbdir = "database/database.duckdb")

# Check if connection worked
conn
```


# Check if connected to database - Check for tables in the database
### Same thing as `.table` in python
```{r}
DBI::dbListTables(conn)
```

# Look at one table
It returns another object- It doesn't know how many rows it has
```{r}
egg_table <- tbl(conn, 
               # Select table
               "Bird_eggs")

personnel_table <- tbl(conn, "Personnel")

Bird_nests_table <- tbl(conn, "Bird_nests")

# Look at tables
egg_table

personnel_table

Bird_nests_table
```
# Dbplyer
## Left-Join
```{r}
nest_eggs_df <- Bird_nests_table %>% 
  left_join(egg_table, by = "Nest_ID") %>% 
  group_by(Observer) %>% 
  summarize(total_eggs = n()) %>%
  collect()

# Print Table
nest_eggs_df

```

# Look at tables before left join
```{r}
Bird_nests_table
personnel_table
```


```{r, message=FALSE}
result <- egg_table %>%
  left_join(Bird_nests_table, by = "Nest_ID") %>%
  group_by(Observer) %>%
  summarize(total_eggs = n()) %>%
  inner_join(personnel_table, by = c("Observer" = "Abbreviation")) %>%
  arrange(desc(total_eggs)) %>%
  head(3) %>%
  select(Name, total_eggs) %>%
  collect()

result

```
# Using Show Query
```{r}
result <- egg_table %>%
  left_join(Bird_nests_table, by = "Nest_ID") %>%
  group_by(Observer) %>%
  summarize(total_eggs = n()) %>%
  inner_join(personnel_table, by = c("Observer" = "Abbreviation")) %>%
  arrange(desc(total_eggs)) %>%
  head(3) %>%
  select(Name, total_eggs) %>%
  show_query()
```


# Questions
- `dbplyr` Did not create a temporary table from the query, it managed to do everything in one query. Which is pretty long.

- Yes the query set a limit to 3 rows using SQL LIMIT Clause.


# Closing Connnection
Important to do when finishing
```{r}
DBI::dbDisconnect(conn, shutdown = TRUE)
```

